---
title: "Assignment06_submission"
author: "Bill Henderson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(quantmod) 
library(tseries)
library(forecast)
library(readxl) 
library(fpp2)
library(vars)
library(lmtest)
library(ggplot2)
library(ggfortify)
library(ggpubr)
```

# Part A. Running Three Models: VAR Level, VAR Growth and ARIMA
## Import the following data via quantmod library

```{r}
getSymbols("DCOILWTICO", src="FRED") # Oil price
getSymbols("NASDAQCOM", src="FRED") # NASDAQ stock price
getSymbols("PAYEMS", src="FRED") # Nonfarm payroll jobs
getSymbols("CPILFESL", src="FRED") # Core Consumer Price Index
getSymbols("DFF", src="FRED") # Federal Fund Rate
```
## Set the full sample period from 1986-02-01 to 2025-1-31 and convert the whole series to
monthly frequency. Note: use na.rm=T if there are missing values.
```{r}
date.range = "1986-02-01/2025-01-31"
oil.price = apply.monthly(DCOILWTICO[date.range], colMeans, na.rm = T)
stock = apply.monthly(NASDAQCOM[date.range], colMeans, na.rm = T)
payroll = apply.monthly(PAYEMS[date.range], colMeans, na.rm = T)
idx = apply.monthly(CPILFESL[date.range], colMeans, na.rm = T)
fund.rate = apply.monthly(DFF[date.range], colMeans, na.rm = T)
```

## Plot the series and convert them to the time series object using “ts” function.
```{r}
oil.price.ts = ts(oil.price, frequency = 12, start = c(1986,2))
oil.plot = autoplot(oil.price.ts, main="Oil Price")

stock.ts = ts(stock, frequency = 12, start = c(1986,2))
stock.plot = autoplot(stock.ts, main="NASDAQ Stock Price")

payroll.ts = ts(payroll, frequency = 12, start = c(1986,2))
payroll.plot = autoplot(payroll.ts, main="Payroll")

idx.ts = ts(idx, frequency = 12, start = c(1986,2))
idx.plot = autoplot(idx.ts, main="Consumer Price Index")

fund.rate.ts = ts(fund.rate, frequency = 12, start = c(1986,2))
fund.plot = autoplot(fund.rate.ts, main="Federal Fund Rate")

ggarrange(oil.plot, stock.plot, payroll.plot, idx.plot, fund.plot, 
          nrow = 2, ncol = 3)
```

## Create a separate time series data frame

```{r}
data0 = cbind(oil.price.ts, stock.ts, payroll.ts, idx.ts, fund.rate.ts)
colnames(data0) = c("oil", "stock", "job", "cpi", "frr")

data1 = cbind(
  diff(log(oil.price.ts)), 
  diff(log(stock.ts)), 
  diff(log(payroll.ts)), 
  diff(log(idx.ts)), 
  fund.rate.ts
  )

data1 = data1[complete.cases(data1),]

colnames(data1) = c("oil_g", "stock_g", "job_g", "cpi_g", "frr")
```

## Create the train sets for the above two data frames using
```{r}
data00 = window(data0, start=c(1986, 3), end=c(2024, 11))
```

## Run two VAR models
```{r}

var00 = VAR(data0, p=12, type="both")  
summary(var00)

imp00.oil = irf(var00, impulse="oil", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp00.oil)

imp00.stock = irf(var00, impulse="stock", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp00.stock)

imp00.job = irf(var00, impulse="job", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp00.job)

imp00.cpi = irf(var00, impulse="cpi", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp00.cpi)



var01 = VAR(data1, p=12, type="const")  
summary(var01)

imp01.oil = irf(var01, impulse="oil_g", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp01.oil)

imp01.stock = irf(var01, impulse="stock_g", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp01.stock)

imp01.job = irf(var01, impulse="job_g", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp01.job)

imp01.cpi = irf(var01, impulse="cpi_g", response="frr", n.ahead=12, ortho=F, runs=100)
plot(imp01.cpi)
```

## Plot the forecasts of these two models
```{r}
pred00 = predict(var00, n.ahead = 12, ci=0.95)

pred01 = predict(var01, n.ahead = 12, ci=0.95)

```

