---
title: "Assignment 04"
author: "Bill Henderson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(quantmod)
library(tidyverse)
library(xts)
library(ggplot2)
library(MASS)
library(forecast)
library(quantmod)
library(tseries)
library(urca)
```

# Data Collection and EDA
## Import Data
```{r}
getSymbols("TOTALSA", src="FRED") # Total Vehicle Sales
getSymbols("PAYEMS", src="FRED") # US payroll jobs
getSymbols("DFF", src="FRED") # Federal fund rates
getSymbols("UNRATE", src="FRED") # Unemployment rates
getSymbols("NASDAQCOM", src="FRED") # Stock market prices
getSymbols("USSTHPI", src="FRED") # Housing market prices
getSymbols("CUSR0000SETB01", src="FRED") # Gasoline price in CPI


```
## Visualizations
```{r}
par(mfrow=c(1,1))
plot(TOTALSA)
plot(PAYEMS)
plot(DFF)
plot(UNRATE)
plot(NASDAQCOM)
plot(USSTHPI)
plot(CUSR0000SETB01)

```
## Convert to Quarterly and select the sample period from 1976Q2 to 2023Q2 (June) as the trainset (in-sample)
```{r}
train.period = c("1976-04/2023-04")
test.period = c("2023-06/2024-03")
totalsa.test = apply.quarterly(TOTALSA[test.period], colMeans, na.rm=T)
cusr.train = apply.quarterly(CUSR0000SETB01[train.period], colMeans, na.rm=T)
dff.train = apply.quarterly(DFF[train.period], colMeans, na.rm=T)
nasdaq.train = apply.quarterly(NASDAQCOM[train.period], colMeans, na.rm=T)
payems.train = apply.quarterly(PAYEMS[train.period], colMeans, na.rm=T)
totalsa.train = apply.quarterly(TOTALSA[train.period], colMeans, na.rm=T)
unrate.train = apply.quarterly(UNRATE[train.period], colMeans, na.rm=T)
ussthpi.train = apply.quarterly(USSTHPI[train.period], colMeans, na.rm=T)
```

### Are the datasets stationary
```{r}
plot(cusr.train)
plot(dff.train)
plot(nasdaq.train)
plot(payems.train)
plot(totalsa.train)
plot(unrate.train)
plot(ussthpi.train)

```
Only TOTALSA, DFF, UNRATE, and CUSR are stationary. The others need treatment
```{r}
cusr.train.st = diff(log(cusr.train))
dff.train.st = dff.train
nasdaq.train.st = diff(log(nasdaq.train))
payems.train.st = diff(log(payems.train))
unrate.train.st = unrate.train
ussthpi.train.st = diff(log(ussthpi.train))
totalsa.train.st = totalsa.train

plot(cusr.train.st)
plot(dff.train.st)
plot(nasdaq.train.st)
plot(payems.train.st)
plot(totalsa.train.st)
plot(unrate.train.st)
plot(ussthpi.train.st)
```


# Part A: Structural Model
## Setting up Time Series
```{r}
cusr = ts(cusr.train.st, frequency = 4, start=c(1979,2))
dff = ts(dff.train.st, frequency = 4, start=c(1979,2))
nas = ts(nasdaq.train.st, frequency = 4, start=c(1979,2))
payems = ts(payems.train.st, frequency = 4, start=c(1979,2))
unrate = ts(unrate.train.st, frequency = 4, start=c(1979,2))
ussthpi = ts(ussthpi.train.st, frequency = 4, start=c(1979,2))
totalsa = ts(totalsa.train.st, frequency = 4, start=c(1979,2))



df = as.data.frame(cbind(cusr, dff,nas,payems,unrate,ussthpi,totalsa))
```
## Generating Linear Regression
```{r}
cor(df, use="complete.obs")

mod01 = lm(totalsa ~ . , data=df)
summary(mod01)

mod01.step = stepAIC(mod01, direction="both")
summary(mod01.step)
```
## ARIMA on Significant Vars
```{r}
dff_for = forecast(auto.arima(dff),4)
nas_for = forecast(auto.arima(nas),4)
unrate_for = forecast(auto.arima(unrate),4)
ussthpi_for = forecast(auto.arima(ussthpi),4)

a_tests = data.frame(dff_for$mean, nas_for$mean, unrate_for$mean, ussthpi_for$mean)
colnames(a_tests) = c("dff", "nas", "unrate", "ussthpi")

a_test.pred = predict(mod01.step,a_tests)
a_test.pred

mean(a_test.pred - as.numeric(totalsa.test))
```
# Part B: A Reduced-form Model
```{r}
totalsa.arima = auto.arima(totalsa)
summary(totalsa.arima)
totalsa.forecast = forecast(totalsa.arima, 4)
plot(totalsa.forecast)
accuracy(totalsa.forecast)

mean(as.numeric(totalsa.forecast$mean) - as.numeric(totalsa.test))
```
# Part C. Mixed Form Model
```{r}
lag.mod = lm(totalsa ~ lag(totalsa) + dff + unrate, data = df)
step.coef = as.matrix(lag.mod$coefficients)

test1 = c(1, totalsa[length(totalsa)], dff_for$mean[1], unrate_for$mean[1])
test1.ans = test1 %*% step.coef

test2 = c(1, test1.ans, dff_for$mean[2], unrate_for$mean[2])
test2.ans = test2 %*% step.coef

test3 = c(1, test2.ans, dff_for$mean[3], unrate_for$mean[3])
test3.ans = test3 %*% step.coef

test4 = c(1, test3.ans, dff_for$mean[4], unrate_for$mean[4])
test4.ans = test4 %*% step.coef

mean(test1.ans - as.numeric(totalsa.test[1]),
     test2.ans - as.numeric(totalsa.test[2]),
     test3.ans - as.numeric(totalsa.test[3]),
     test4.ans - as.numeric(totalsa.test[4]),
     )
```


