---
title: "Assignment05_Submission"
author: "Bill Henderson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(lubridate)
library(forecast)
library(MASS)
library(prophet)
```

# Data Import

```{r}
# trainset data is from 4/1/2016 to 2/28/2017 
# testset data is from 3/1/2017 to 3/19/2017
ds = read_excel("TVlift.xlsx") %>%
  mutate(dow = factor(wday(date)),
         month = (month(date)),
         dom = (day(date)))
train = ds[ds$date <= as.Date("2017-02-28"),]
test = ds[ds$date > as.Date("2017-02-28"),]

"Training Set"
  train %>%
    summarize(first_date = min(date), last_date = max(date)) %>%
    kbl() %>%
    kable_styling(full_width = F )

"Test Set"
  test %>%
    summarize(first_date = min(date), last_date = max(date)) %>%
    kbl() %>%
    kable_styling(full_width = F )
```
# Models
## Linear Regressions
### Quick EDA
```{r}
train %>%
  dplyr::select(-date) %>%
  summary() %>% 
  kbl() %>% 
  kable_styling(full_width = F)

cor(dplyr::select(train,-c(date,dom,dow,month))) %>% kbl() %>% kable_styling(full_width = F)

plot(train)


boxplot(train$sb)
boxplot(train$snb)
boxplot(train$tv)
boxplot(train$sales)

## convert to long format with tidyr::pivot_longer
  pivot_longer(train, cols = snb:tv) |>
  ggplot(aes(x=date, y=value, colour=name)) +
  geom_line() 

  
  pivot_longer(train, cols = c(sb,sales)) |>
  ggplot(aes(x=date, y=value, colour=name)) +
  geom_line() 
```
No NA's in our data site, which is very nice. Seeing a very tight correlation between snb and sales. "snb" and tv may be a little too tightly correlated as well, but we'll try and see how it looks.Also appears to have some stationarity and homoskedacity issues with a couple vars.

May be some outliers that could be removed as well, especially above the upper quartile.


## Models
```{r}
reg01 = lm(sales~sb +snb +tv +dow +month +dom, train[train$date >= as.Date(2016-11-10),])
ggAcf(residuals(reg01))
s01 = summary(reg01)

reg02 = lm(sales~sb +snb +tv +dow +month, train[train$date >= as.Date(2016-11-10),])
ggAcf(residuals(reg02))
s02 = summary(reg02)

reg03 = lm(sales~sb +snb +tv +dow +month + lag(sales,1), train[train$date >= as.Date(2016-11-10),])
ggAcf(residuals(reg03))
s03 = summary(reg03)

reg04 = lm(sales~sb +snb +dow +month + lag(sales,1), train)
ggAcf(residuals(reg04))
s04 = summary(reg04)

reg05 = lm(sales~sb +snb +dow +month + lag(sales,1) + lag(tv,7), train)
ggAcf(residuals(reg05))
s05 = summary(reg05)

reg06 = lm(sales~sb +snb +dow +month + lag(sales,1) + lag(tv,14), train)
ggAcf(residuals(reg06))
s06 = summary(reg06)

# removed the highly correlated snb from the model with all sig. vars
# with highest R^2
reg07 = lm(sales~sb +dow +month + lag(sales,1), train)
ggAcf(residuals(reg07))
s07 = summary(reg07)

step = stepAIC(reg01)
ggAcf(residuals(step))
step.sum = summary(step, direction="both")

tv.lim.df = train[train$date >= as.Date(2016-11-10),]
ari01 = auto.arima(tv.lim.df$sales, xreg = cbind(tv.lim.df$sb, tv.lim.df$snb, tv.lim.df$tv))
ggAcf(residuals(ari01))
s.ari01 = summary(ari01)

data.frame(metrics = c("r.sq","adj.r.sq", "std err"),
           reg01 = c(s01$r.squared, s01$adj.r.squared, s01$sigma),
           reg02 = c(s02$r.squared, s02$adj.r.squared, s02$sigma),
           reg03 = c(s03$r.squared, s03$adj.r.squared, s03$sigma),
           reg04 = c(s04$r.squared, s04$adj.r.squared, s04$sigma),
           reg05 = c(s05$r.squared, s05$adj.r.squared, s05$sigma),
           reg06 = c(s06$r.squared, s06$adj.r.squared, s06$sigma),
           reg07 = c(s07$r.squared, s07$adj.r.squared, s07$sigma),
           step = c(step.sum$r.squared, step.sum$adj.r.squared, step.sum$sigma)
           ) %>%
  mutate(across(2:9, \(x) round(x,3))) %>%
  kbl() %>%
  kable_styling(full_width = F)

s.ari01
```
## Predictions

```{r}
reg.preds = data.frame(
  date = test$date,
  sales.actual = test$sales
  ) %>%
  mutate(
    pred.ari01 = predict(ari01, newxreg = cbind(test$sb, test$snb, test$tv))$pred,
    pred01 = predict(reg01, test),
    pred02 = predict(reg02, test),
    pred03 = predict(reg03, test),
    pred04 = predict(reg04, test),
    pred05 = predict(reg05, test),
    pred06 = predict(reg06, test),
    pred07 = predict(reg07, test),
    step.pred = predict(step, test)
  )

rmse01 = sqrt(mean((reg.preds$pred01-reg.preds$sales.actual)^2, na.rm = T))
rmse02 = sqrt(mean((reg.preds$pred02-reg.preds$sales.actual)^2, na.rm = T))
rmse03 = sqrt(mean((reg.preds$pred03-reg.preds$sales.actual)^2, na.rm = T))
rmse04 = sqrt(mean((reg.preds$pred04-reg.preds$sales.actual)^2, na.rm = T))
rmse05 = sqrt(mean((reg.preds$pred05-reg.preds$sales.actual)^2, na.rm = T))
rmse06 = sqrt(mean((reg.preds$pred06-reg.preds$sales.actual)^2, na.rm = T))
rmse07 = sqrt(mean((reg.preds$pred07-reg.preds$sales.actual)^2, na.rm = T))
rmse.ari = sqrt(mean((reg.preds$pred.ari01-reg.preds$sales.actual)^2, na.rm = T))
rmse.step = sqrt(mean((reg.preds$step.pred-reg.preds$sales.actual)^2, na.rm = T))

data.frame(metrics = c("rmse"),
           reg01 = rmse01,
           reg02 = rmse02,
           reg03 = rmse03,
           reg04 = rmse04,
           reg05 = rmse05,
           reg06 = rmse06,
           reg07 = rmse07,
           reg.arima = rmse.ari,
           reg.step = rmse.step) %>%
  mutate(across(2:10, \(x) round(x,3))) %>%
  kbl() %>%
  kable_styling(full_width = F)

```

```{r}
## convert to long format with tidyr::pivot_longer
  pivot_longer(reg.preds, cols = c(sales.actual, pred03, pred04, pred05, pred.ari01)) |>
  ggplot(aes(x=date, y=value, colour=name)) +
  geom_line() 

s05
```


